---
title: "Butzke_HWK_LDA_9"
author: "Alden Butzke"
date: "5/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Linear Discriminant Analysis 

We will analyze loans to determine whether they should be approved or rejected based on  
1.liquidity,  
2.profitability,  
3.activity.  
Thus, we will develop a rule through LDA that will sort loan applications into accept groups and reject groups with corresponding clusterheads. 


```{r}
setwd("/Users/aldenbutzke/Desktop/Datamining/")
library(readxl)
library(lattice)
library(caret)
rm(list=ls())

df<-read.csv("Loans.csv", header = T)
```


```{r}
df$col<-ifelse(df$Category==1, "red", "blue")

x1L<- mean(df$Liquidity[df$Category==1])

x1P<-mean(df$Profitability[df$Category==1])

x1A<-mean(df$Activity[df$Category==1])

```

```{r}
LiqMeans<-tapply(df$Liquidity, # what column 
                  
                  df$Category, # how to group 
                  
                  mean #function to apply
                  
                  )

ProfMeans<-tapply(df$Profitability, df$Category, mean)

ActMeans<-tapply(df$Activity, df$Category, mean)

Clusterheads<-data.frame(Group=c(1,2), Liquidity=LiqMeans, Profitability=ProfMeans, Activity=ActMeans)

```


$Group =\beta_0 + \beta_1 X_1 + \beta_2 X_2+ \beta_3 X_3$ 

```{r}
LDA<-lm(Category~Liquidity+Profitability+Activity, data = df)
sLDA<-summary(LDA)
sLDA
```

```{r}
df$Predict<-4.3071 - 2.8554*df$Liquidity - 0.9880*df$Profitability - 0.2540*df$Activity

PredM<-predict(LDA, newdata = Clusterheads)
Clusterheads$PredM<-PredM
hist(df$Predict, breaks = 50)
```
  
### The Model 

The clusterhead for group 1 is 1.26 and the clusterhead for group 2 is 1.74. 

The Cutoff is 1.5. 
```{r}
cutoff<-mean(PredM)
cutoff
df$Predict.Final<-ifelse(df$Predict<=cutoff,1,2)
```
```{r}
out<-table(df$Category, df$Predict.Final)
out
confusionMatrix(out)

```

### Accuracy  


The model is 81.63% accurate. The accuracy was sourced by generating a confusion matrix from the LDA sorting of predicted groups by clusterheads on profitability, activity, and liquidity.  

```{r}
new.data<-readxl::read_xlsx("New.Data.xlsx")



new.data$Predict<-4.3071 - 2.8554*new.data$Liquidity - 0.9880*new.data$Profitability - 0.2540*new.data$Activity

new.data$Predict.Final<-ifelse(new.data$Predict<=cutoff,1,2)

new.data

```

### New Data

You should accept A and B and reject C, D and E because our model predicts with near 80% accuracy that A and B closely fit in with group 1 and C, D and E closely fit with group 2

