---
title: "Butzke_HW8"
author: "Alden Butzke"
date: "4/19/2022"
output: html_document
---


### Linear Regression 2

Loading in the data and adding dummy variables based on quarter in order to estimate the following equation:  

$\hat{y}=β_0+β_1 x+β_2 d_1+β_3 d_2+β_4 d_3+ϵ $

where $d_i$ variables are dummy variables for the quarters. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
getwd()
setwd("/Users/aldenbutzke/Desktop/Datamining/")
getwd()

rm(list=ls())

library(readxl)
library(tidyverse)

x<-read_excel("Retail.xlsx")
head(x)
library(fastDummies)
x<-dummy_cols(x, select_columns = "Quarter")
head(x)

```
### Creating a linear model 

```{r}
mod1<-lm(data = x, `Retail Sales`~GNP + Quarter_1 + Quarter_2 + Quarter_3 )
ms<-summary(mod1)
c1<-round(coefficients(mod1),2)

print(paste0("The Equation is, y=", c1[1],"B0 ", "+", c1[2],"B1 ", "+", c1[3],"B2 ", "+", c1[4],"B3 ", "+", c1[5],"B4 "))

Q2_predict<- 263995.074443642+ 13000*53.133109960926-0*98646.0803104984-26350.6178189479-0*38084.9289368957

Q4_predict<- 263995.074443642+ 13000*53.133109960926-0*98646.0803104984-0*26350.6178189479-0*38084.9289368957

```

The predicted retail sales in Q2 with GNP = 13,000 billion are $`r Q2_predict` billion 

The predicted retail sales in Q4 with GNP = 13,000 billion are $`r Q4_predict` billion 

### Hypothesis Testing 

Testing the following hypothesis to determine whether Q1-3 = Q4  

H_0:β_0=0 vs H_A:β_0≠0

We will test this with $ \alpha $ = 0.05, the 5% significance level. 

```{r}
p<-ms$coefficients[1,4]
```

we find that `r p` < $ \alpha $ = 0.05, thus we can reject the null that β_0=0. 