---
title: "Butzke_ICE2"
author: "Alden Butzke"
date: "5/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/Users/aldenbutzke/Desktop/Datamining/")
library(readxl)
library(lattice)
library(tidyverse)
library(caret)
rm(list=ls())

```

```{r}
df<-read.csv("StoreMixUp.csv", header = T)
Index<-c(1:length((df$Pofits_100s)))
df$Index<-Index
df1<-data.frame(df$Num_Sales, df$Pofits_100s, df$Index)
plot(df$Num_Sales, df$Pofits_100s)
#plot(df1$Num_Sales,df1$Pofits_100s)
#x<-is.na(df1$df.Num_Sales)
#sum(x)
            
            

```

### Group analysis 

There is clear evidence for multiple groups because there are visually obvious clusters with distinct proft to number of sales ratios. There are at least 6 visually obvious groups. 

```{r}
hist(df$Pofits_100s, breaks = 200)
hist(df$Num_Sales, breaks = 200)
```

The histograms are multi-modal, there is further evidence of multiple groups. 

```{r}
library(cluster)
library(factoextra)
library(FactoMineR)
library(fBasics)
```

```{r}

NAdf <- is.na.data.frame(df1)
any(NAdf)
set.seed(27500)
k<-kmeans(df1[,c(1,2)], centers=6, nstart = 10)
df$k<-k$cluster
df1$k<-k$cluster




ggplot(df1, aes(df.Num_Sales, df.Pofits_100s))+
  geom_point(aes(col=k))

```
###  Problems 

The outliers of each group are being miscategorized. Thus, I will create a boxplot for each group in order to house each observation with its actual group.  

```{r}
#g1<- boxplot(df$Pofits_100s[df$k==1], 
            # horizontal = T)
#g1

fviz_cluster(k, geom = "point", data = df)
```

We can see clearly that the data has been better fitted through a process of trial and error to accurately group the observations with a new algorithim. 


```{r}
df$Col<-ifelse(df$k == 1, "black",
  ifelse(df$k == 2, "red", 
   ifelse(df$k == 3, "green",
    ifelse(df$k == 4, "blue",
     ifelse(df$k == 5, "lightblue", "purple")))))

ggplot(df, aes(Num_Sales, Pofits_100s))+
  geom_point(aes(col=Col))

B1<-df[df$k==1,]
R2<-df[df$k==2,]
G3<-df[df$k==3,]
B4<- df[df$k==4,]
LB5<- df[df$k==5,]
P6<-df[df$k==6,]

boxplot(df$Pofits_100s~df$k,
        main= "Clusters and Profit",
        col= df$Col,
        ylab= "Profits",
        xlab = "Cluster")

  boxplot(df$Num_Sales~df$k,
        main= "Clusters and Profit",
        col= df$Col,
        ylab= "Sales",
        xlab = "Cluster")

```
```{r}
P1<-plot(df$Pofits_100s[df$k==1]~df$Num_Sales[df$k==1], 
             main= "Black 1",
             col= "black")

P2<-plot(df$Pofits_100s[df$k==2]~df$Num_Sales[df$k==2], 
           main= "Red 2",
           col= "red")

P3<-plot(df$Pofits_100s[df$k==3]~df$Num_Sales[df$k==3],
             main = "Green 3",
             col= "green"
             )

P4<-plot(df$Pofits_100s[df$k==4]~df$Num_Sales[df$k==4],
             main = "Blue 4",
             col= "blue"
             )
P5<-plot(df$Pofits_100s[df$k==5]~df$Num_Sales[df$k==5],
             main = "Light Blue 5",
             col= "light blue"
             )
P6<- plot(df$Pofits_100s[df$k==6]~df$Num_Sales[df$k==6],
             main = "Purple 6",
             col= "purple"
             )
```

```{r}
B1O<-B1[B1$Num_Sales>500,]
B1<-B1[B1$Num_Sales<500,]

Cluster1<- rbind(R2, B1O) 

R2O<-R2[R2$Pofits_100s>300,]
R2<-R2[R2$Pofits_100s<300,] 

Cluster2<- rbind(R2, R2O) 

G3O<-G3[G3$Num_Sales>130,]
G3<-G3[G3$Num_Sales<130,] 

Cluster3<- rbind(B1O, G3O) 


B4O<-B4[B4$Pofits_100s>120,]
B4<-B4[B4$Pofits_100s<120,] 

Cluster4<- rbind(G3, B4O)

LB5O<-LB5[LB5$Num_Sales<50,]
LB5<-LB5[LB5$Num_Sales>50,] 

Cluster5<- rbind(G3, LB5O)

P6O<-P6[P6$Pofits_100s>350,]
P6<-P6[P6$Pofits_100s<350,]


Cluster6<- rbind(P6, R2O)

fdf<-rbind(Cluster1,Cluster2,Cluster3,Cluster4,Cluster5,Cluster6)

plot(fdf$Pofits_100s~fdf$Num_Sales, col= fdf$Col, ylab ="Profit (100s)", xlab= "Number of Sales")
```
The groups have been sorted into discrete bunches and though it isn't perfect, it is a big improvement. I graphed each group and looked for clear outliers and eyeballed where those breaks occured. I could have been more precise by calculating mathmatical outliers using boxplots and reassinging colors automatically. However, the data had clear visual splits. 

```{r}
mb1<-mean(B1$Pofits_100s)
mr2<-mean(R2$Pofits_100s)
mg3<-mean(G3$Pofits_100s)
mb4<-mean(B4$Pofits_100s)
mlb5<-mean(LB5$Pofits_100s)
mp6<-mean(P6$Pofits_100s)


mdf<-data.frame(mb1,mr2,mg3,mb4,mlb5,mp6)
#plot(mdf)

plot(c(mb1, mr2, mg3, mb4, mlb5, mp6), xlab="groups", ylab="profit means")


mb1<-mean(B1$Num_Sales)
mr2<-mean(R2$Num_Sales)
mg3<-mean(G3$Num_Sales)
mb4<-mean(B4$Num_Sales)
mlb5<-mean(LB5$Num_Sales)
mp6<-mean(P6$Num_Sales)


mdf<-data.frame(mb1,mr2,mg3,mb4,mlb5,mp6)

plot(c(mb1, mr2, mg3, mb4, mlb5, mp6), xlab="groups", ylab="Sales Num Means")
#xplot(mdf)
```
This project nearly killed me and I wish I was in class when we went over it, but I want to learn a lot more about k means and unsurpervised learning. 
