---
title: "Choosing.Rgeression.Model"
author: "Alden Butzke"
date: "4/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Choosing Regression Model

```{r}
getwd()
setwd("/Users/aldenbutzke/Desktop/Datamining")
rm(list=ls())
```


**Preliminaries**
newline in RMD = two spaces then enter.  

We will load the data first.

#### Read in Data:
To create a new code chunk, we can do two things,  
1) hit Ctrl+Alt+i  
2) hit apple button + cmmnd + i
```{r}
College<-read.csv("College.csv", header = T)
### visualize all numerical data, can use pairs 
pairs(College[2:6])
```

### Choosing the best model
It is not an optimum, it is the best we can find given the data (historical data)

*Two main facts*  

1. $R^2$ always increases as you add more variables to the model  

2. When a model has more than two variables, you must use Adjusted$R^2$  

**Check Simple Relationships**  
You are going to develop linear models for each variable vs dependent
Our Y is Earnings.  

1. Earnings versus Cost: $\hat{y}=\beta_0+\beta_1*x_1$  
2. Earnings versus Grad Rates:  $\hat{y}=\beta_0+\beta_1*x_2$ 
3. Earnings versus Debt:  $\hat{y}=\beta_0+\beta_1*x_3$ 
4. Earnings versus City:  $\hat{y}=\beta_0+\beta_1*x_4$ 

```{r}
## run simple regression models 
#attach(College)
#detach(College) # not working
lm1<- lm(College$Earnings~College$Cost)
lm2<- lm(College$Earnings~College$Grad)
lm3<- lm(College$Earnings~College$Debt)
lm4<- lm(College$Earnings~College$City)

s1<-summary(lm1)
s2<-summary(lm2)
s3<-summary(lm3)
s4<-summary(lm4)

```

How to pull info from the coe into text:  

1. Model 1, $x_1$ = Cost, the $R^2$ value is, `r s1$r.squared`
2. Model 2, $x_2$ = Grad, the $R^2$ value is, `r s2$r.squared`
3. Model 3, $x_3$ = Debt, the $R^2$ value is, `r s3$r.squared`
4. Model 4, $x_4$ = City, the $R^2$ value is, `r s4$r.squared`  

The most explanatory variable to explain the variation in Y~Earnings is Graduation Rates  

```{r}
int2<-round(s2$coefficients[1], 2)
sl2<-round(s2$coefficients[2], 2)
```

Given by the model, $\hat{y_i}=$$`r int2`+`r sl2`$$*Grad$ with an $R^2=$ `r s2$r.squared`. 

now we check 2 variable relationships 
```{r}
lm21<-lm(data = College, Earnings~Grad+Cost)
s21<-summary(lm21)
lm22<-lm(data = College, Earnings~Grad+Debt)
s22<-summary(lm22)
lm23<-lm(data = College, Earnings~Grad+City)
s23<-summary(lm23)
```


The adjusted R squared are:  

1. `r round(s21$adj.r.squared, 3)`  
2. `r round(s22$adj.r.squared, 3)`  
3. `r round(s23$adj.r.squared, 3)`  

```{r}
#The best model so far is model 1, $y_i=$ `r round(s21$coefficients[1], 2)` + `r round(`
lm21

```

```{r}
lm221<-lm(data = College, Earnings~Grad+Cost+Debt)
s221<-summary(lm221)
lm232<-lm(data = College, Earnings~Grad+Cost+City)
s232<-summary(lm232)
```

1. `r round(s221$adj.r.squared, 3)`  
2. `r round(s232$adj.r.squared, 3)` 


```{r}
## run a model with all data 

df<-College[2:6]
All<-lm(data=df, Earnings~.)
All
all.s<-summary(All)
```

1. `r round(all.s$adj.r.squared, 3)`  

