---
title: "Histogram and Density Plots"
author: "Alden Butzke"
date: "3/8/2022"
output: html_document
---

### Histograms and Density Plots 

headers are defined by the number of hashtags or pound symbols 

ctrl+alt+i will create a new code block 
```{r}
rm(list=ls())
College<-read.csv("College.csv", header = TRUE)
#min for cost and earnings columns 
lowest<-min(College[,2:3])
highest<-max(College[,2:3])



```

lowest is `r lowest` 

```{r}
p1<-hist(College$Earnings,  #plot earnings 
         breaks = 20, #will generate 20 bins 
         col = rgb(1,0,0,1/2)
         )

p2<-hist(College$Cost, breaks = 10, col = rgb(0,1,0,1/4))
```

```{r}
plot(p1, 
     col = rgb(0,0,1,1/4), 
     xlim=c(9000,75000),
     main = "Histograms of Earnings and Cost", 
     xlab = "Dollars ($)")

plot(p2, col = rgb(1,0,0,1/4), 
     xlim = c(9000,75000), add = TRUE
     
     
     
     )






```

can run chi squred, duncan difference, regression, or one other test to see if data is normal 

### Density Block

1. all probability p(X) for each variable is a number between 0 and 1 
2. all probability p(X) added together results in 1 

```{r}
d1<-density(College$Grad)
d2<-density(College$Debt)

plot(d1,
     xlim=(c(20,100)),
     ylim=(c(0,.09)),
     lty="longdash",
     lwd=2,
     col="blue4",
     main = "Density plots of Grad and Debt Rates"
     )
lines(d2, col="red", lwd=1.5, lty="twodash")
legend(20,.06, #location of legend
       legend = c("Grad_Rate", "Debt-Rate"), #labels for legend 
       col = c("blue4", "red"),
       lty = c("longdash", "twodash")
       )
```

### Standardization

z-score is the most common method used to standardize. $z_i=\frac{x_i-\mu}{\sigma}$


```{r}

###  z-score is the most common method used to standardize. $z_i=\frac{x_i-\mu}{\sigma}$

#### logs linearize data that is sepeerated by multiple orders of magnitude 

m<-mean(College$Debt) #mean
stdd<-sd(College$Debt) #calc sd 
newDebt<-((College$Debt-m)/stdd)

hist(newDebt, breaks = 10, xlim= c(-6,6), col = "blue4", main = "Histograms using a manual Z score")

#### There is a function for standardization called scale 
stdGrad<-scale(College$Grad, center=TRUE, scale = TRUE)
hist(stdGrad, breaks = 10, xlim= c(-6,6), col = "blue4", main = "Histograms using scale")

```
 

```{r}
d1<-density(scale(College$Debt))
d2<-density(scale(College$Grad))
plot(d1, xlim=c(-6,6), ylim=c(0,0.5), col="blue4", lty="longdash")
lines(d2, col="red", lty="twodash") 
legend(-6,.5,legend = c("Debt", "Grad"), col = c("blue4", "red"), lty = c("longdash", "twodash"))

boxplot(College[,4:5])


```

### Dirty Data

```{r}
pairs(College[,2:5])
```

