---
title: "OrdinalCat.Nominal"
author: "Alden Butzke"
date: "4/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
getwd()
setwd("/Users/aldenbutzke/Desktop/Datamining/")
getwd()

library(readxl)
rm(list=ls())

list.files()

SM<-read_excel("Supermarket", sheet = "Data")

head(SM)
```

### Different data types

```{r}
a<-length(unique(SM$`Customer ID`))

print(paste0("The NUmer of Unique Customers is: ", a ))

length(SM$`Customer ID`[])
```

Factors are a method to convert nominal categories into numerical data. Factorization is the ability to assign a unique number to a category type at level. The level is the number of possible catgeories a variable can take on.  

Gender=M, assign a 1, gender+F, assign a 0  

```{r}
gender<-as.factor(SM$Gender)
```
Level of new column

```{r}
levels(gender)
```
Levels function only works on columns/variables that are in factor form 

#### Rules of level 

levels makes the data 1 and 2, but we want it to be 0 and 1. THus, we can make an overfit model and set the levels by using the function. Set the level for F = 0 and M=1.  

ALWAYS leave a comment on the conversion 

```{r}
levels(gender)<-c(0,1)
Revenue<- SM$Revenue

mod1<-lm(Revenue~gender)
s1<-summary(mod1)
s1

plot(Revenue~gender)
abline(mod1)

```

```{r}
df<-SM[,c("Revenue", "Gender")]
aggregate(df$Revenue, by= list(df$Gender), FUN=mean)


```

### Dummy Variable Creation 

We will create a column that serves as an indicator column that has 1 for the case being present and 0 otherwise

```{r}
df$Fem<-ifelse(df$Gender== "F",1,0)
df$Mal<-ifelse(df$Fem=="1",0,1)


mod2<-lm(data=df, Revenue~Fem+Mal)
s2<-summary(mod2)
s2
```


```{r}

married<-as.factor(SM$`Marital Status`)
levels(married)<-c(0,1) # setting married = 0, single = 1 
mod3<-lm(Revenue~married) 
s3<-summary(mod3)

plot(Revenue~married)
```


### Categorical  Data with more than 2 levels 

```{r}
SM$`Annual Income`[1:4]
head(SM$`Annual Income`)
inc<-table(SM$`Annual Income`)
#inc1<-table(SM$`Annual Income`[["$150K +"]])
#inc1 

length(unique(SM$`Annual Income`))
```

```{r}
library(fastDummies)
datadum<-dummy_cols(SM, select_columns = "Annual Income")
head(datadum)
max(datadum$`Annual Income_$150K +`)
```
```{r}
#mod4<-lm(datadum, Revenue~datadum[,17,18,19,20])
#plot_ly(x = gender, y = Revenue)
library(plotly)
l <- htmltools::tagList()
for (i in 1:3) {
  l[[i]] <- plot_ly(x = rnorm(10))
}
l
```

