---
title: "Outliers"
author: "Alden Butzke"
date: "3/3/2022"
output: powerpoint_presentation
---

## Outlier Detection

- We have a college data set of 113 small colleges throughout the US.
- The data consists of:
  + Name of College
  + Graduation rate within 4 years
  + Debt Rate
  + Earnings after 5 years
  + Cost after 4 years
  
## Brief look at data by using code box
```{r echo=FALSE}
College<-read.csv("College.csv", header = TRUE)
boxplot(College[,c(2,3)],
        col = c("Chocolate", "Orange"), 
        horizontal = TRUE, border = "Black", notch = TRUE, 
        xlab="Money ($)")
```
  
  
## Outliers Detected in Earnings

- We will use the limits of:

    
```{r echo=FALSE}
EarnSummary<-summary(College$Earnings)
Q1=as.numeric(EarnSummary[2])
Q3=as.numeric(EarnSummary[5])
IQR=Q3-Q1
LL=as.numeric(Q1-(1.5*IQR))
UL=as.numeric(Q3+(1.5*IQR))

```
  + Lower limit: $LL=Q_1-1.5*IQR$=`r LL`
  + Upper Limit: $UL=Q_3+1.5*IQR$=`r UL`
  
```{r echo=FALSE}
Outliers<-College$School[College$Earnings>UL]
OutEarn<- College$Earnings[College$Earnings>UL]
Outliers1<-Outliers[1]

```

## Brief View at Graduation and Debt Rates

```{r echo=FALSE}
College<-read.csv("College.csv", header = TRUE)
boxplot(College[,c(4,5)],
        col = c("Chocolate", "Orange"), 
        horizontal = TRUE, border = "Black", notch = TRUE, 
        xlab="Percentage Rate %")
### Debt 
SD<-summary(College$Deb)
Q1D<- as.numeric(SD[2])
Q3D<- as.numeric(SD[5])
IQRD<-Q3D-Q1D
LLD<- Q1D-1.5*IQRD
ULD<- Q3D+1.5*IQRD

OutliersD<- College$School[College$Debt<LLD]
```


## Outliers in Debt Rates

- 4 Outliers found with low debt rates.
- These outliers are:
```{r echo=FALSE}
print(paste0(OutliersD))

```


## Outliers in Grad Rates
- 1 Oulier Detected with a low graduation rate
```{r echo=FALSE}
SG<-summary(College$Grad)
Q1G<-as.numeric(SG[2])
Q3G<-as.numeric(SG[5])
IQRG<-Q3G-Q1G
LLG<-Q1G-1.5*IQRG

OutliersG<-College$School[College$Grad<LLG]

```

- The Outlier is:
```{r echo=FALSE}
print(paste0(OutliersG))
```

---

